Getting and Cleaning Data - Assignment 1
========================================================

## Path variables
Define all the file locations where X_<set>.txt are the measurements, y_<set>.txt are the activities and subject_<set>.txt are the subjects ID.
```{r echo=TRUE}
# file locations
test.data.path="test/X_test.txt"
test.activity.path="test/y_test.txt"
test.subject.path="test/subject_test.txt"
train.data.path="train/X_train.txt"
train.activity.path="train/y_train.txt"
train.subject.path="train/subject_train.txt"
activity.label.path="activity_labels.txt"
```


## Util functions
Create a function is.mean.or.std to identify if the column is a mean or std or not. If it is a numeric string is returned, NULL otherwise.
```{r echo=TRUE}
# Filter all measurments with mean or std only
is.mean.or.std <- function(column){
        if (grepl(".*mean\\(\\)|.*std\\(\\)",column)){
                return ("numeric")
        }else{
                return("NULL")
        }
}
```

Create a function to 
1. read only field required (only mean/std in our case, definded by features input).
2. add the subject column (which is read from the file)
3. add the activity column (which is read from the file)
4. merge with activity map to get proper naming of an activity
- datapath: raw data to be read
- activitypath: file location containing each activity of each row
- subjectpath: file location containing subject of each row
- features: dataframe with "numeric","NULL" and name of measurement feature
- activitymap: dataframe of ID/activity name mapping
```{r echo=TRUE}
# This function will read data of the fpath and only features required
# (to save memory)
# assume 
read.only.required <- function(datapath, activitypath, subjectpath, features, activitymap){
        
        # Load Data
        data <- read.table(datapath, header=F, colClasses=features$required, col.names=features$Name)
        
        # Add Subject column
        data.subject <- read.table(subjectpath, header=F, col.names=c("SubjectID"))
        data$SubjectID <- data.subject$SubjectID
        
        # Add Activity column
        data.activity <- read.table(activitypath, header=F, col.names=c("ActivityID"))
        data$ActivityID <- data.activity$ActivityID
        
        # Using merge to get activity ID -> activity Name
        merged.data <- merge(data,activity.map, all=TRUE)
        
        # Return the final data set
        merged.data
}
```


## Create data frame for data read
Create the feature data frame such that 
feature$require is either numeric or NULL
feature$Name is the name of each measurement feature

```{r echo=TRUE}
# Prepare the features mapping, only get those means and stds
features <- read.table("features.txt", header=F, as.is=T, col.names=c("ID", "Name"))
features$required <- sapply(features$Name, FUN=is.mean.or.std)
```

Create activity map such that
activity.map$ActivityID is the id of each activity
activity.map$Activity is the name of each activity

```{r echo=TRUE}
# Prepare the activity mapping
activity.map <- read.table(activity.label.path, header=F, col.names=c("ActivityID", "Activity"))
activity.map$Activity <- as.factor(activity.map$Activity)
```


## Load Test and train data and merge them
```{r echo=TRUE}
# get decorated test data
test.data <- read.only.required(test.data.path, 
                                test.activity.path,
                                test.subject.path,
                                features,
                                activity.map)


# get decorated train data
train.data <- read.only.required(train.data.path, 
                                 train.activity.path,
                                 train.subject.path,
                                 features,
                                 activity.map)
        
# Append test data/train data
merged.data <- rbind(test.data, train.data)
```


## Calculate average of each variable for each activity and each subject.
```{r echo=TRUE}
# Calculate average of each variable for each activity and each subject. 
library(data.table)
my.data.table <- data.table(merged.data)
my.mean.data <- my.data.table[, lapply(.SD, mean), by=c("SubjectID", "Activity")]
my.mean.data <- my.mean.data[order(my.mean.data$SubjectID),]
```


## Export the data and save it as "tidy.data.mean.txt"
```{r echo=TRUE}
write.table(my.mean.data, "tidy.data.mean.txt", sep="\t")
```

